---
title: "Meta Analysis"
author: "Rachel Donahue, Yuli Jin, Xiang Li, Boyu Chen, Zhihui Zhang"
date: "12/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=F,message = F , echo = T)
knitr::opts_chunk$set(fig.width=8, fig.height=4,fig.align = "center")
library(tidyverse)
library(meta)
library(metafor)
library(openxlsx)
library(stringr)
library(grid)
```

This is a report showing the results of a regenerated meta analysis on studies investigating the effects of dietary interventions on hypertension in Sub-Saharan Africa. The first table shows the studies and the data structure that fueled the underlying analysis, and included in this report are funnel plots with interpretations showing that there does not appear to be any publication bias, forest plots, bubble plots and a meta regression. The meta regression was fit with one predictor of study duration, and it has a small affect on SBP and DBP but is not significant.

## Data Processing Results

```{r echo=FALSE}
data<-read.xlsx("Final Studies_Meta Analysis_1.3.xlsx",startRow = 2)
meta <- data[,c(1,3,6:8,10:12)]
names(meta) <- c("StudyID","Author" ,"Size", "MeanDiff_SBP", "MeanDiff_DBP", "SE_SBP", "SE_DBP", "Duration_M")

# add study ID and Author to 2nd 3rd row
meta[2:3,1:2]=meta[1,1:2]
# add duration
meta[c(2,3,6),'Duration_M']=c('2','2','6')
# input nan in SE_SBP and SE_DBP
tmp_sbp<-str_extract_all(meta$SE_SBP,"-*\\d+.\\d+")%>% map(`[`,2) %>% unlist() %>%  na.omit() 
meta$SE_SBP[c(2,7)]=tmp_sbp
tmp_dbp<-str_extract_all(meta$SE_DBP,"-*\\d+.\\d+")%>% map(`[`,2) %>% unlist() %>%  na.omit() 
meta$SE_DBP[c(2,7)]=tmp_dbp
# input SE study ID 1(placebo)
meta[3,c('SE_SBP','SE_DBP')]=c('3.004','2.593')

# manually change Mean_Diff_SBP& DBP of study IDs 1&5:
meta[1,c('MeanDiff_DBP')]=c('-1.8')
meta[2,c('MeanDiff_SBP','MeanDiff_DBP')]=c('-11.2','-4.9')
meta[3,c('MeanDiff_SBP')]=c('-1.0')
meta[7,c('MeanDiff_SBP','MeanDiff_DBP')]=c('-2.54','-3.95')


#manually change the Duration of study 5:
meta[7,c('Duration_M')]=c('6')

# extract the first number and omit the content after \n
meta<-meta %>% mutate_at(c('MeanDiff_SBP','MeanDiff_DBP','SE_SBP','SE_DBP'),~str_extract(.,"-*\\d+.\\d+") %>% as.numeric(.))
# drop na columns
meta<-meta[c(!is.na(meta$SE_DBP)),]
# delete et al
meta$Author<-str_extract(meta$Author,'\\w+')
# adjust the mean of study ID 6 into -4.8 and -3.2
meta[8,c('MeanDiff_SBP','MeanDiff_DBP')]=-1*meta[8,c('MeanDiff_SBP','MeanDiff_DBP')]
# remove placebo group the third Siervo study
meta<-meta[-3,]

```
```{r echo=FALSE}
knitr::kable(meta,'pipe')

```



## Forest & Funnel Plots

```{r echo=FALSE}
rownames(meta)<-seq(1:dim(meta)[1])
meta$Author=paste(rownames(meta),meta$Author,sep='.')
# ?metagen
# ?funnel.meta
meta_SBP <- metagen(TE = MeanDiff_SBP, seTE = SE_SBP, studlab = Author,data = meta, sm = "MD", comb.fixed = F, comb.random = T, method.tau = "SJ", title = "SBP", prediction = TRUE)

```


```{r echo=FALSE}
forest_SBP <- forest.meta(meta_SBP, layout = "RevMan5")
grid.text("SBP Forest Plot", .5, .9, gp=gpar(cex=1.5))
```
The forest plot shows that the mean difference for SBP is generally negative, the highest weighted study is Study ID 6, Forrester in terms of it's effects on the meta analysis.



```{r label='SBP_forest_plot_store', eval=FALSE,echo=FALSE}
# this chunk is used to store jpeg
jpeg('graph/SBP_forest_plot.jpeg',width = 2200, height = 1100, res=300)
forest_SBP <- forest.meta(meta_SBP, layout = "RevMan5")
grid.text("SBP Forest Plot", .5, .9, gp=gpar(cex=1.5))
dev.off()
```




```{r echo=FALSE}
meta_SBP <- metagen(TE = MeanDiff_SBP, seTE = SE_SBP, studlab = Author,data = meta, sm = "MD", comb.fixed = F, comb.random = T, method.tau = "SJ", title = "SBP", prediction = TRUE)
funnel.meta(meta_SBP,studlab = TRUE,cex.studlab = 1,cex = 1,pch=16,pos.studlab=c(4,1,2,1,1,1,1,1),xlim=c(-15,15))
title(main = "SBP Funnel Plot",  cex.main = 1.5,font.main = 1)

```






```{r label='SBP_funnel_plot_store',eval=FALSE,echo=FALSE}
# this chunk is used to store jpeg
jpeg('graph/SBP_funnel_plot.jpeg',width = 2200, height = 1100, res=300)
meta_SBP <- metagen(TE = MeanDiff_SBP, seTE = SE_SBP, studlab = Author,data = meta, sm = "MD", comb.fixed = F, comb.random = T, method.tau = "SJ", title = "SBP", prediction = TRUE)
funnel.meta(meta_SBP,studlab = TRUE,cex.studlab = 1,cex = 1,pch=16,pos.studlab=c(4,1,2,1,1,1,1,1),xlim=c(-15,15))
title(main = "SBP Funnel Plot",  cex.main = 1.5,font.main = 1)
dev.off()
```


This is a fairly balanced funnel plot with an even amount of studies on either side of the mean difference line, and the studies relatively follow the expected funnel shape. The mean difference line is a negative value, which makes sense given that the studies are involving interventions to reduce SBP, it would be unlikely that there would be a study published with a positive mean difference. Overall this plot shows that there does not appear to be any publication bias.


```{r echo=FALSE}

meta_DBP <- metagen(TE = MeanDiff_DBP, seTE = SE_DBP, studlab = Author, data = meta, sm = "MD", comb.fixed = F, comb.random = T, method.tau = "SJ", title = "DBP", prediction = TRUE)
forest_DBP <- forest.meta(meta_DBP, layout = "RevMan5")
grid.text("DBP Forest Plot", .5, .9, gp=gpar(cex=1.5))
```

This is a forest plot for DBP mean difference, which is also generally negative, and the highest weighted studies are Schouw and Forrester.

```{r echo=FALSE}
funnel.meta(meta_DBP,  studlab = TRUE,cex.studlab = 1,cex = 1,pch=16,pos.studlab=c(4,1,2,2,1,1,1,1),xlim=c(-6,6))
title(main = "DBP Funnel Plot",  cex.main = 1.5,font.main = 1)
```


```{r label='DBP_store',echo=FALSE,eval=FALSE}
# this chunk is used to store jpeg
jpeg('graph/DBP_forest_plot.jpeg',width = 2200, height = 1100, res=300)
meta_DBP <- metagen(TE = MeanDiff_DBP, seTE = SE_DBP, studlab = Author, data = meta, sm = "MD", comb.fixed = F, comb.random = T, method.tau = "SJ", title = "DBP", prediction = TRUE)
forest_DBP <- forest.meta(meta_DBP, layout = "RevMan5")
grid.text("DBP Forest Plot", .5, .9, gp=gpar(cex=1.5))
dev.off()

jpeg('graph/DBP_funnel_plot.jpeg',width = 2200, height = 1100, res=300)
funnel.meta(meta_DBP,  studlab = TRUE,cex.studlab = 1,cex = 1,pch=16,pos.studlab=c(4,1,2,2,1,1,1,1),xlim=c(-6,6))
title(main = "DBP Funnel Plot",  cex.main = 1.5,font.main = 1)
dev.off()

```

Similar to SBP, this is a fairly balanced funnel plot with an even amount of studies on either side of the mean difference line, and the studies relatively follow the expected funnel shape. The mean difference line is a little higher but still negative value, it seems in one study there was a positive mean difference in DBP after intervention. Overall this plot indicates that there does not appear to be any publication bias.

## Meta regression

```{r echo=FALSE}

M<-meta$Duration_M%>% as.numeric()
duration <- M*30
duration_m <- M - 1
duration_m2 <- M - 2
duration_lg <- log(duration_m)
size <- meta$Size

```
```{r echo=FALSE}
# SBP
meta_SBP_reg <- metareg(meta_SBP, ~duration_m2)
meta_SBP_reg

bubble(meta_SBP_reg, studlab = TRUE,xlim=c(-6,25),ylim = c(-12,0),pos.studlab=c(2,2,2,2,1,2,2))




```
The coefficient for duration is negative, showing that as the number of days in a study increases, the more DBP decreases on average.The I squared value is at 48.46%, indicating a moderate amount of heterogeneity, indicating that the variation in study outcomes is less likely due to chance.



```{r echo=FALSE}
# DBP
meta_DBP_reg <- metareg(meta_DBP, ~duration_m2)
meta_DBP_reg

bubble(meta_DBP_reg, studlab = TRUE,xlim=c(-6,25),ylim = c(-6,0),pos.studlab=c(1,2,2,4,2,2,2))


```


```{r label='bubble_store',echo=FALSE,eval=FALSE}
# this chunk is used to store jpeg
jpeg('graph/SBP_bubble_plot.jpeg',width = 2200, height = 1100, res=300)
bubble(meta_SBP_reg, studlab = TRUE,xlim=c(-6,25),ylim = c(-12,0),pos.studlab=c(2,2,2,2,1,2,2))
dev.off()

jpeg('graph/DBP_bubble_plot.jpeg',width = 2200, height = 1100, res=300)
bubble(meta_DBP_reg, studlab = TRUE,xlim=c(-6,25),ylim = c(-6,0),pos.studlab=c(1,2,2,4,2,2,2))
dev.off()
```
The coefficient for duration is negative, showing that as the number of days in a study increases, the more DBP decreases on average.The I squared value is at 32.91%, indicating a smaller amount of heterogeneity, that the variation in study outcomes is more likely due to chance.


