---
title: "Meta Analysis"
author: "Rachel Donahue, Yuli Jin, Xiang Li, Boyu Chen, Zhihui Zhang"
date: "12/10/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=F,message = F , echo = T)
knitr::opts_chunk$set(fig.width=8, fig.height=4,fig.align = "center")
library(tidyverse)
library(meta)
library(metafor)
library(openxlsx)
library(stringr)
library(grid)
```

## Data Processing Results

```{r echo=FALSE}
data<-read.xlsx("Final Studies_Meta Analysis_1.3.xlsx",startRow = 2)
meta <- data[,c(1,3,6:8,10:12)]
names(meta) <- c("StudyID","Author" ,"Size", "MeanDiff_SBP", "MeanDiff_DBP", "SE_SBP", "SE_DBP", "Duration_M")

# add study ID and Author to 2nd 3rd row
meta[2:3,1:2]=meta[1,1:2]
# add duration
meta[c(2,3,6),'Duration_M']=c('2','2','6')
# input nan in SE_SBP and SE_DBP
tmp_sbp<-str_extract_all(meta$SE_SBP,"-*\\d+.\\d+")%>% map(`[`,2) %>% unlist() %>%  na.omit() 
meta$SE_SBP[c(2,7)]=tmp_sbp
tmp_dbp<-str_extract_all(meta$SE_DBP,"-*\\d+.\\d+")%>% map(`[`,2) %>% unlist() %>%  na.omit() 
meta$SE_DBP[c(2,7)]=tmp_dbp
# input SE study ID 1(placebo)
meta[3,c('SE_SBP','SE_DBP')]=c('3.004','2.593')

# manually change Mean_Diff_SBP& DBP of study IDs 1&5:
meta[1,c('MeanDiff_DBP')]=c('-1.8')
meta[2,c('MeanDiff_SBP','MeanDiff_DBP')]=c('-11.2','-4.9')
meta[3,c('MeanDiff_SBP')]=c('-1.0')
meta[7,c('MeanDiff_SBP','MeanDiff_DBP')]=c('-2.54','-3.95')


#manually change the Duration of study 5:
meta[7,c('Duration_M')]=c('6')

# extract the first number and omit the content after \n
meta<-meta %>% mutate_at(c('MeanDiff_SBP','MeanDiff_DBP','SE_SBP','SE_DBP'),~str_extract(.,"-*\\d+.\\d+") %>% as.numeric(.))
# drop na columns
meta<-meta[c(!is.na(meta$SE_DBP)),]
# delete et al
meta$Author<-str_extract(meta$Author,'\\w+')
# adjust the mean of study ID 6 into -4.8 and -3.2
meta[8,c('MeanDiff_SBP','MeanDiff_DBP')]=-1*meta[8,c('MeanDiff_SBP','MeanDiff_DBP')]
# remove placebo group the third Siervo study
meta<-meta[-3,]

```
```{r echo=FALSE}
knitr::kable(meta,'pipe')

```



## Forest & Funnel Plots

```{r echo=FALSE}
rownames(meta)<-seq(1:dim(meta)[1])
meta$Author=paste(rownames(meta),meta$Author,sep='.')
# ?metagen
# ?funnel.meta
meta_SBP <- metagen(TE = MeanDiff_SBP, seTE = SE_SBP, studlab = Author,data = meta, sm = "MD", comb.fixed = F, comb.random = T, method.tau = "SJ", title = "SBP", prediction = TRUE)

```


```{r echo=FALSE}
forest_SBP <- forest.meta(meta_SBP, layout = "RevMan5")
grid.text("SBP Forest Plot", .5, .9, gp=gpar(cex=1.5))

```

```{r echo=FALSE}
meta_SBP <- metagen(TE = MeanDiff_SBP, seTE = SE_SBP, studlab = Author,data = meta, sm = "MD", comb.fixed = F, comb.random = T, method.tau = "SJ", title = "SBP", prediction = TRUE)
funnel.meta(meta_SBP,studlab = TRUE,cex.studlab = 1,cex = 1,pch=16,pos.studlab=c(4,1,2,1,1,1,1,1),xlim=c(-15,15))
title(main = "SBP Funnel Plot",  cex.main = 1.5,font.main = 1)

```

This is a fairly balanced funnel plot with an even amount of studies on either side of the mean difference line, and the studies relatively follow the expected funnel shape. The mean difference line is a negative value, which makes sense given that the studies are involving interventions to reduce SBP, it would be unlikely that there would be a study published with a positive mean difference. Overall this plot shows that there does not appear to be any publication bias.


```{r echo=FALSE}

meta_DBP <- metagen(TE = MeanDiff_DBP, seTE = SE_DBP, studlab = Author, data = meta, sm = "MD", comb.fixed = F, comb.random = T, method.tau = "SJ", title = "DBP", prediction = TRUE)
forest_DBP <- forest.meta(meta_DBP, layout = "RevMan5")
grid.text("DBP Forest Plot", .5, .9, gp=gpar(cex=1.5))
funnel.meta(meta_DBP,  studlab = TRUE,cex.studlab = 1,cex = 1,pch=16,pos.studlab=c(4,1,2,2,1,1,1,1),xlim=c(-6,6))
title(main = "DBP Funnel Plot",  cex.main = 1.5,font.main = 1)
```
Similar to SBP, this is a fairly balanced funnel plot with an even amount of studies on either side of the mean difference line, and the studies relatively follow the expected funnel shape. The mean difference line is a little higher but still negative value, it seems in one study there was a positive mean difference in DBP after intervention. Overall this plot indicates that there does not appear to be any publication bias.

## Meta regression

```{r echo=FALSE}

M<-meta$Duration_M%>% as.numeric()
duration <- M*30
duration_m <- M - 1
duration_m2 <- M - 2
duration_lg <- log(duration_m)
size <- meta$Size

# SBP
meta_SBP_reg <- metareg(meta_SBP, ~duration_m2)
meta_SBP_reg

bubble(meta_SBP_reg, studlab = TRUE)

# DBP
meta_DBP_reg <- metareg(meta_DBP, ~duration_m2)
meta_DBP_reg

bubble(meta_DBP_reg, studlab = TRUE)


```

